Collision rates of planetesimals near mean-motion resonances

### 2020-10-30 refereeResponse.txt

35) pg.6 lcol P.1 l.33 - "cylindrical distance" - please, define before start using it.

    A sentence has been added to the text to clarify what cylindrical distance means.

36) pg.6 rcol P.2 l.47-54 - "This also suggests... radial collision profile." - I don't know if I agree with authors' conclusions that MMRs are not that important, and that most collisions occur with objects out of resonance. It looks for me that the features seem in Fig 6 (similar to Fig 4) are caused by objects in MMRs colliding at their pericentric and apocentric distances within the eccentric orbits. In my understanding, if authors' arguments are correct, MMRs should basically not matter, and Fig 5 would look similar if no planet was present. If that is the case, how can this work serve to constrain the mass and eccentricity of a perturbing planet?

    As shown in figure 5, the bump and dip features at the center of the MMRs are still present, even if we ignore particles whose semimajor axis falls within the libration width of the resonances. This does not mean, however, that the resonances are not important. As we approach the edges of the resonance from the outside, the critical angle approaches a stationary value. Where the critical angle is stationary, the pericenter orientations of the planetesimals experience a torque which causes their \varpi to follow the giant planet in its orbit. This is what we propose is responsible for the coherent structure seen in the right-hand panels of figures 6 and 7, along with the structure seen in the collision profiles. TODO: Expand our explanation in the paper to clarify this.

@@

37) pg.6 rcol P.2 l.38-56 - "As mentioned in section 2, ... appears to 'follow' the planet." - This paragraph is super confuse. I don't understand this. Why not simply say this is caused by the secular forcing in eccentricity? This would save 3 paragraphs in the paper and would be much more clear.

    See 36). If you look at figures 1, 6 and 7, the nonaxisymmetric gap structure in the disk near the resonances is always aligned with the planet's orbital phase. It is also apparent from figure 1 that the gaps are not aligned with the planet's longitude of pericenter, which should be the case if the structure were due to secular forcing in eccentricity. Furthermore, bodies in resonance experience strong precession, which randomizes the orientations of their orbits. The gap structure cannot be due to the interface between the secularly forced non-resonant part of the disk, and the randomly oriented pericenters of bodies in resonance. The latter explanation was described in Tabeshian & Weigery 2016, which we do not find consistent.

@@

38) Figs 4 and 5 - Why is the relative # larger than 1? Relative to what?

    Figures 4 and 5 are showing the probability density of the collision rates. The curves on these plots are normalized such that the area under the curve = 1. For the orange curve in figure 5, the normalization factor is scaled by the ratio between the total number of collisions in each sample. Some text has been added alongside the discussion of figures 4 and 5 to explain this.

@@

42) pg.12 rcol P.1 l.8-24 - "Near the interior 2:1... both of these quantities." - This is a clear way to set the explanation. Well done! I hope to see more of that in the revised version. I think a lot of the explanations in the main text should be substituted by something clear and concise as this! Make it simple!

    :-)

43) pg.12 rcol P.3 l.45-48 - "Given that the solid surface density diminishes with distance (r^-1.5 for the MMSN (Hayashi 1981)), exterior resonances may not be quite as useful for this type of measurement, however.)" - Not sure I agree. Exterior MMRs can more easily capture inward drifting objects (e.g. Weidenschilling & Davis 1985; Patterson 1987; Beauge & Ferraz-Mello 1993; Malhotra 1993; Gomes 1995; Jiang & Yeh 2004; Batigyn & Laughlin 2015; Storch & Batygin 2019), increasing their local collision rates (Deienno et al. 2020), thus potentially generating very strong dusty features.

    Absolutely true if we are going to consider radial drift. Given that an expanded discussion of radial drift is now an important part of this paper, we should be more careful not to dismiss the importance of exterior MMRs. TODO: How to change the text to address this?

### 2020-10-30 main.tex
More edits, add figure to clarify gravitational focusing explanation

Next, we examine the statistics of collisions resolved in each of the simulations. The 3D positions and velocities of the two colliding bodies are 
recorded to a table at the moment of impact. We derive the Keplerian orbital elements of a collision using these positions and velocities. 
First, we examine the semimajor axis of the first collider, which is shown in figure \ref{fig:coll_hist_a}. During a collision, the ``first'' particle is defined 
as the more massive of the two. However, nearly all of the collisions happen between the initial, equal-mass planetesimals. In this case, the 
distinction is set by the collision search algorithm and is rather arbitrary. In all of the plots where we show collision statistics, we have verified that 
using the ``first'' or ``second'' collider does not qualitatively change any of the features. We find that some of the features present in figure 
\ref{fig:coll_hist_a} in this and subsequent figures are highly sensitive to the number of bins and the location of the bin edges. For this reason, we 
construct a \textbf{probability density function (PDF)} of the collisions using a Kernel Density Estimate (KDE). We use the {\sc neighbors.KernelDensity} function from the {\sc sklearn} 
\citep{scikit-learn} package to construct our KDEs, using a gaussian kernel with a FWHM of 0.02 au.

@@

We attribute the production of a bump or a dip at certain semi-major axes to changes in the gravitational focusing factor during two-body encounters. 
Closer to the star, the planetesimal disc starts off dynamically colder. Here, additional heating from the resonance acts to suppress the gravitational 
focusing cross section, which reduces the collision rate from its
non-perturbed value. Further out in the disc, where the unperturbed
encounter velocity is higher, the gravitational focusing cross section
is already minimized. Any additional heating instead acts to simply
increase the encounter rate between bodies, which enhances the
collision rate. This explains the relative drop in the collision rate
near the 3:1 MMR and the relative increase near the 2:1 and 5:3
resonances. Although this could potentially serve as a useful
diagnostic of the planetesimal size in a disc (because the mutual
escape velocity sets how dynamically hot the bodies can get before
gravitational focusing is suppressed), a direct measurement of the
semimajor axes of the colliding planetesimals is not
possible. Furthermore, as we will demonstrate next, collisions due to bodies in resonance do not have a significant effect on the final shape of the radial dust distribution.

@@

@@

Operating under this assumption\footnote{As discussed in \cite{2017ApJ...850..103B}, the local generation of second-generation dust is only one 
consideration.  The dust will also evolve spatially due to drag effects, but can also evolve in size or be re-accreted onto planetesimals. For simplicity, 
we focus on local dust sourcing only while acknowledging that subsequent evolution could affect the morphology.}, a map of the relative concentration 
of second-generation dust can be constructed from the cylindrical distance at which the collisions occur. \textbf{Here, cylindrical distance is measured as the distance between the central star and the collision in the ($r, \theta$) plane.} This is shown in figure \ref{fig:coll_hist_r}. 
Similar to figure \ref{fig:coll_hist_a}, we use a KDE with a width of 0.02 au to assemble the collision locations into a radial distribution. Most 
strikingly, the bump that was present near the 2:1 resonance is no longer visible in cylindrical distance space, although it suddenly appears again in 
the e3m2 simulation. The dip that is visible near the 3:1 MMR also becomes a bump in the highest eccentricity case. To determine how the resonant 
bodies actually contribute to the radial collision profile, we excluded collisions that fall between $2.495 < a < 2.505$ au (near the 3:1 resonance) and 
between $3.2 < a < 3.35$ au (near the 2:1 resonance), which is shown by the dashed curve. Qualitatively, none of the bump or dip features present 
are removed by making this exclusion. This suggests that the features near resonance seen in semimajor axis space become smeared out in 
cylindrical distance space due to the large spread in eccentricity and orbital orientation.

>>>>

Next, we examine the statistics of collisions resolved in each of the simulations. The 3D positions and velocities of the two colliding bodies are 
recorded to a table at the moment of impact. We derive the Keplerian orbital elements of a collision using these positions and velocities. 
First, we examine the semimajor axis of the first collider, which is shown in figure \ref{fig:coll_hist_a}. During a collision, the ``first'' particle is defined 
as the more massive of the two. However, nearly all of the collisions happen between the initial, equal-mass planetesimals. In this case, the 
distinction is set by the collision search algorithm and is rather arbitrary. In all of the plots where we show collision statistics, we have verified that 
using the ``first'' or ``second'' collider does not qualitatively change any of the features. We find that some of the features present in figure 
\ref{fig:coll_hist_a} in this and subsequent figures are highly sensitive to the number of bins and the location of the bin edges. For this reason, we 
construct a \textbf{probability density function (PDF)} of the collisions using a Kernel Density Estimate (KDE). We use the {\sc neighbors.KernelDensity} function from the {\sc sklearn} 
\citep{scikit-learn} package to construct our KDEs, using a gaussian kernel with a FWHM of 0.02 au. \textbf{The curves shown in figure \ref{fig:coll_hist_a} are normalized such that the area underneath is normalized to 1.}

@@

We attribute the production of a bump or a dip at certain semi-major axes to changes in the gravitational focusing factor during two-body encounters.
\textbf{Rewrite this paragraph to reference figure \ref{fig:gf}.}
Closer to the star, the planetesimal disc starts off dynamically colder. Here, additional heating from the resonance acts to suppress the gravitational 
focusing cross section, which reduces the collision rate from its
non-perturbed value. Further out in the disc, where the unperturbed
encounter velocity is higher, the gravitational focusing cross section
is already minimized. Any additional heating instead acts to simply
increase the encounter rate between bodies, which enhances the
collision rate. This explains the relative drop in the collision rate
near the 3:1 MMR and the relative increase near the 2:1 and 5:3
resonances. Although this could potentially serve as a useful
diagnostic of the planetesimal size in a disc (because the mutual
escape velocity sets how dynamically hot the bodies can get before
gravitational focusing is suppressed), a direct measurement of the
semimajor axes of the colliding planetesimals is not
possible. Furthermore, as we will demonstrate next, collisions due to bodies in resonance do not have a significant effect on the final shape of the radial dust distribution.

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/gf_plot.png}
    \caption{\textbf{The collision rate of bodies in the vicinity of the 3:1 (left) and 2:1 (right) MMRs, relative to the minimum value, as a function of the local eccentricity dispersion. The pairs of points connected by lines show the values of the perturbed and unperturbed collision rates in the e1m2, e2m2 and e3m2 simulations.}\label{fig:gf}}
\end{center}
\end{figure*}

@@

Operating under this assumption\footnote{As discussed in \cite{2017ApJ...850..103B}, the local generation of second-generation dust is only one 
consideration.  The dust will also evolve spatially due to drag effects, but can also evolve in size or be re-accreted onto planetesimals. For simplicity, 
we focus on local dust sourcing only while acknowledging that subsequent evolution could affect the morphology.}, a map of the relative concentration 
of second-generation dust can be constructed from the cylindrical distance at which the collisions occur. \textbf{Here, cylindrical distance is measured as the distance between the central star and the collision in the ($r, \theta$) plane.} This is shown in figure \ref{fig:coll_hist_r}. 
Similar to figure \ref{fig:coll_hist_a}, we use a KDE with a width of 0.02 au to assemble the collision locations into a radial distribution. Most 
strikingly, the bump that was present near the 2:1 resonance is no longer visible in cylindrical distance space, although it suddenly appears again in 
the e3m2 simulation. The dip that is visible near the 3:1 MMR also becomes a bump in the highest eccentricity case. To determine how the resonant 
bodies actually contribute to the radial collision profile, we excluded collisions that fall between $2.495 < a < 2.505$ au (near the 3:1 resonance) and 
between $3.2 < a < 3.35$ au (near the 2:1 resonance), which is shown by the dashed curve. \textbf{As in figure \ref{fig:coll_hist_a}, the solid curves are normalized such that the area underneath is equal to 1. For the dashed curves, the normalization factor is scaled down according to the total number of collisions in the entire sample and the number of collisions in the subsample which excludes the collisions in resonance.} Qualitatively, none of the bump or dip features present 
are removed by making this exclusion. This suggests that the features near resonance seen in semimajor axis space become smeared out in 
cylindrical distance space due to the large spread in eccentricity and orbital orientation.
